---
title: "Obtaining Population Data"
author: "Andrew Deighan"
date: "July 5, 2017"
output: html_document
---

```{r, echo=FALSE, cache=TRUE}
downloaddate <- date()
```

  The population data was obtained from the National Cancer Institute's Surveillance, Epidemiology, and End Results (SEER) program https://seer.cancer.gov/popdata/download.html. This website has census bureau county level population data (by age group) going back to 1969. We accessed this data on `r downloaddate`

```{r downloaddata, results="hide", cache=TRUE}
library(readr)

url <- "https://seer.cancer.gov/popdata/yr1969_2015.19ages/us.1969_2015.19ages.txt.gz"
file <- tempfile(pattern = "populationdata", fileext = ".gz")
download.file(url, file)

popdata <- read_fwf(file,
                     fwf_positions(c(1,5,19),
                                   c(4,6,26),
                                   c("year", "state", "population")),
                     skip = 3001703)

unlink(file)
```

<br>

  Now we will tidy up the data a little more, including changing the state name abbreviations to full state names (this will make it easier when we eventually merge this data set with our other data sets)
  
```{r cleanup, results="hide"}
library(dplyr)

popdata <- popdata %>% 
  filter(year >= 1980) %>%
  mutate(population = as.integer(population)) %>%
  group_by(year, state) %>% 
  summarise(population = sum(population)) %>%
  ungroup()

snames <- tolower(state.name)
for(r in 1:nrow(popdata)){
    if(length(snames[state.abb == popdata$state[r]]) > 0){
    popdata$state[r] <- snames[state.abb == popdata$state[r]]
    }
}

popdata$state[popdata$state == "DC"] <- "district of columbia"

write.csv(popdata, "data/Population Data.csv", row.names = FALSE)
```